---
title: "WUP_04_Alt"
author: "Connor Rice"
date: "09/12/2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
gssdata <- read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and the RACE of the person taking the test   
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two different RACES

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(gssdata,aes(x=RACE,y=WORDSUM))

basicCN + geom_boxplot()
```

The graph shows only very small differences between the White and Black races scores, with identical medians and nearly identical 50th percentile ranges. Only their mins, maxs, and 75th percentiles differ significantly. However both the White and Black races data is noticeably higher on average with their medians being nearly above the 75th percentile of the Other race(s).

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~RACE,data=gssdata)
```

There is very little difference in the means of the White and Black races, however both the mean of the Whtie race and the Black race is over a full point higher than the mean of the category of Other race(s).

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~RACE,data=gssdata)
summary.aov(model1)
```

The p-value of 0.0182 shows their is a significant difference between the three categories of data with regard to race. This disproves the null hypothesis but does not detail where the difference lies.

```{r}
plot(TukeyHSD(model1),las=2)
```

The final plot indicates that only the difference between the White race and the Other race(s) is significant since its 95% confidence range does not cover 0. There is likely no significant difference between either the Black and Other race(s) nor the Black and White race(s) due to their 95% confidence ranges both falling on/around 0.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

The overall data shows a small significant difference between the White and Other race(s) but no significant difference between the other categories. However both the Black and Other race(s) have massively smaller sample sizes than the White race which makes comparison between any of the categories difficult to interpret with any significant level of confidence.
